---
title: "Dealing with Bad Column Names"
author: "Matt Pettis (matthew.pettis@gmail.com)"
date: "May 23, 2018"
output:
  html_document:
    code_folding: show
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

One of the nice things about `R` is that columns from a data frame can be referenced like regular code names (or variables) within `R` code.  This makes it much easier to work on data columns at a time with plain `R` syntax.  However, `R` code names have particular restrictions in how they are named, such as having to start with a dot or an alphabetic character, and contain only alphanumeric characters, along with dots and underscores.  Column names from data objects like CSV files, do not have such a restriction.  In order to use column names like code names, it is helpful to have standard ways to convert CSV names, with more relaxed conventions, to `R` names that follow `R` conventions.

The `janitor` package can convert column names, among other tasks, to be `R`-usable column names.  Furthermore, CSV columns can have column names repeated more than once, but names in an `R` dataframe must have unique names, and `janitor` appends numerical suffixes to make unique names when it needs to.

Here we show how to do that.

# Code

Here we make some data with non-`R` compliant code names, and repeat the last field:

```{r}
# Setup ------------------------------------------------------------------------
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(purrrlyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(rprojroot))

suppressPackageStartupMessages(library(janitor))


# Get project root directory
dir_proj <- find_root(has_file("PROJECT_ROOT"))



# Create data ------------------------------------------------------------------
set.seed(123)
dat <- tibble(
    cat = sample(c("red", "green", "blue"), 22, replace = TRUE)
  , val = sin(2 * pi * seq_along(cat) / 20) + rnorm(length(dt), sd = 0.05)
  , val2 = sin(2 * pi * seq_along(cat) / 20) + rnorm(length(dt), sd = 0.05)
)

# Make names that are problematic
names_prob <- c(
  "Acceleration-group\n(kg-m/s^2)"
  , "Bayes % prob\t[breaking confidence #]"
  , "Bayes % prob\t[breaking confidence #]"
)

# Put those names on the dataframe
# dat <- set_names(dat, names_prob)
dat <- setNames(dat, names_prob)

glimpse(dat)
```

You can see that the column names here have a lot of non alpha-numeric characters, like spaces, brackets, newlines, and tabs to name a few.  We can use the `janitor::clean_names()` function to make `R` compliant names.  But it is always a good habit to store away the old names in case they are needed down the rowd.

```{r}
# Store away original column names
colnames_orig <- names(dat)

# Make new, R-usable column names
dat <- dat %>%
  clean_names()

# Save original headers for reference, but use as the name of this vector the new, transformed names:
colnames_orig <- set_names(colnames_orig, names(dat))
colnames_orig %>% print()

# Access old name via new name, example
colnames_orig["bayes_percent_prob_breaking_confidence_number_2"]
```
