---
title: "Inconsistent Naming Conventions CSV column names with non-standard characters"
author: "Matt Pettis (matthew.pettis@gmail.com)"
date: "June 19, 2018"
output:
  html_document:
    code_folding: show
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

It is easy and common in Excel to make cells (like column headers) that have embedded newlines, because it makes column headers look nice.  But it wreaks havoc on using those column names as references within code.  Here, we show you how that can be fixed.



# Data creation

Make data with inconsistent naming conventions.

```{r}
# Setup ------------------------------------------------------------------------
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(purrrlyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(rprojroot))
suppressPackageStartupMessages(library(janitor))


# Get project root directory
dir_proj <- find_root(has_file("PROJECT_ROOT"))


# Create data ------------------------------------------------------------------
set.seed(123)
dat <- tibble(
    cat = sample(c("red", "green", "blue"), 22, replace = TRUE)
  , val = sin(2 * pi * seq_along(cat) / 20) + rnorm(length(dt), sd = 0.05)
  , val2 = sin(2 * pi * seq_along(cat) / 20) + rnorm(length(dt), sd = 0.05)
)

# Make names that are problematic
names_prob <- c(
  "Acceleration-group\n(kg-m/s^2)"
  , "Bayes % prob\t[breaking confidence #]"
  , "Bayes % prob\t[breaking confidence #]"
)

# Put those names on the dataframe
# dat <- set_names(dat, names_prob)
dat <- setNames(dat, names_prob)

glimpse(dat)
```

This data has two types of problems -- it's got embedded problematic characters, and a repeated column header.





# Solution

You can see that the column names here have a lot of non alpha-numeric characters, like spaces, brackets, newlines, and tabs to name a few.  We can use the `janitor::clean_names()` function to make `R` compliant names.  But it is always a good habit to store away the old names in case they are needed down the road.


```{r}
# Store away original column names
colnames_orig <- names(dat)

# Make new, R-usable column names
dat <- dat %>%
  clean_names()

# Save original headers for reference, but use as the name of this vector the new, transformed names:
colnames_orig <- set_names(colnames_orig, names(dat))
colnames_orig %>% print()

# Access old name via new name, example
colnames_orig["bayes_percent_prob_breaking_confidence_number_2"]
```
