---
title: "Inconsistent Naming Conventions"
author: "Matt Pettis (matthew.pettis@gmail.com)"
date: "June 19, 2018"
output:
  html_document:
    code_folding: show
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

Sometimes you will receive files that should contain monthly data.  It would be nice if those filenames followed a pattern.  Often, they don't.  This is just one example of how to rename those files.

For a good reference on how to make good and useful naming conventions, see here: http://www2.stat.duke.edu/~rcs46/lectures_2015/01-markdown-git/slides/naming-slides/naming-slides.pdf



# Data creation

Make data with inconsistent naming conventions.

```{r}
# Setup ------------------------------------------------------------------------
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(purrrlyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(rprojroot))
# suppressPackageStartupMessages(library(janitor))


# Get project root directory
dir_proj <- find_root(has_file("PROJECT_ROOT"))



# Create data ------------------------------------------------------------------
set.seed(123)

dat <- tibble(
    dt = seq(from=ymd("2018-01-01"), to=ymd("2018-05-31"), by="day")
  , val = sin(2 * pi * seq_along(dt) / 7) + rnorm(length(dt), sd = 0.05)
)

# See the data
print(dat)

# Write data to local directory ------------------------------------------------

# Setup directory to write to
dir_dat <- file.path(dir_proj, "dat", "inconsistent_naming_conventions")
if (!dir.exists(dir_dat)) dir.create(dir_dat, recursive = TRUE)

# Write out individual files
write_csv(dat %>% filter(format(dt, "%Y-%m") == "2018-01"), file.path(dir_dat, "Jan18.csv"))
write_csv(dat %>% filter(format(dt, "%Y-%m") == "2018-02"), file.path(dir_dat, "2018-February.csv"))
write_csv(dat %>% filter(format(dt, "%Y-%m") == "2018-03"), file.path(dir_dat, "2018-03.csv"))
write_csv(dat %>% filter(format(dt, "%Y-%m") == "2018-04"), file.path(dir_dat, "The-Cruelest-Month.csv"))
write_csv(dat %>% filter(format(dt, "%Y-%m") == "2018-05"), file.path(dir_dat, "04 (May) 2018.csv"))
```






# Solution

This is not really recommended in general, it is best to leave names as you see them.  But if you really want to, you can rename them.  I will do this if I don't expect more data, and I expect that I will pass on data to other people, and this is a way of being nice.

```{r}
# Read in the file names
input_file_fullpaths <- list.files(dir_dat, full.names = TRUE)

# Process each file.
# Read in each file, figure out which month it belongs to, make a new consistent name
# for file, and write out the file.
input_file_fullpaths %>%
  walk(function(input_filepath) {
    # Read in file
    datl <- read_csv(input_filepath, col_types = cols())
    
    # Detect what month it is from
    year_mon_str <- datl$dt %>%
      format("%Y-%m") %>%
      unique()
    
    # Write out file with naming convention of "<yyyy-mm>.csv"
    write_csv(datl, file.path(dir_dat, sprintf("%s.csv", year_mon_str)))
  })
```
