---
title: "Time series with missing entries"
author: "Matt Pettis (matthew.pettis@gmail.com)"
date: "June 19, 2018"
output:
  html_document:
    code_folding: show
    number_sections: yes
    theme: paper
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Introduction

Time series data is common.  Many models require that time series data be equally spaced.  If you use models like that, you should figure out if your data is actually complete and equally spaced, as well as non-duplicated.  If it isn't, can you pinpoint where?



# Data creation

```{r}
# Setup ------------------------------------------------------------------------
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(purrrlyr))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(rprojroot))
# suppressPackageStartupMessages(library(janitor))


# Get project root directory
dir_proj <- find_root(has_file("PROJECT_ROOT"))


# Create data ------------------------------------------------------------------
set.seed(123)

# Original data
dat <- tibble(
    dt = seq(from=ymd("2018-01-01"), to=ymd("2018-05-31"), by="day")
  , val = sin(2 * pi * seq_along(dt) / 7) + rnorm(length(dt), sd = 0.05)
)

# Duplicate some rows
dat <- bind_rows(dat, dat[28:34,])

# Delete a run of rows
dat <- dat[-(77:88),]
```






# Solution

Easist thing to do here is make an explicit series of the timerange in the interval that you need, and left join a count of your observations by time onto that.

```{r}
# Make a summary table, counting each row 
dat_freq <- dat %>%
  count(dt)

dat_freq %>% print()

# Create *all* expected time stamps
full_date_seq <- seq(from=min(dat_freq$dt), to=max(dat_freq$dt), by="day")

# Take full time range, left join on data we have
dat_full <- tibble(dt=full_date_seq) %>%
  left_join(dat_freq, "dt")

# Plot the dates
ggplot(dat_full, aes(dt, n)) +
  geom_col(width = 0.8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)
        , strip.text.y = element_text(angle = 0)) +
  scale_x_date(date_labels = "%Y-%m-%d", date_breaks = "1 week") +
  ggtitle("dt row frequency in data")
```
